---
title: "Simple D3 Example"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

D3 is a popular Javascript library for visualizations. It can also be used to select and change node elements in a webpage. Additionally, there is an R package, [R2D3](https://rstudio.github.io/r2d3/) that allows for D3 code to be run easily in Shiny.

This workflow was adapted from an `HTML` implementation of D3 on an old website I made. I wanted to reuse some of the code because I like the effect that it makes and adapt it to run in `R` and `Shiny`. This is my first usage of the R2D3 package.

### Load Libraries

```{r}
library(shiny)
library(r2d3)
```

### Create Simple Shiny App

This app won't do anything, hence the empty server function, other than change the ui elements with D3. The JS code is captured in a seperate script that is shown below.

```{r}
ui <- fluidPage(
  r2d3(script = "d3.js", height = 0),# important to set the height of the D3 script
  
  # create test divs
  tags$div(style = "background-color: red; height: 400px",
           id = "one",
           "Initial Div"),
  tags$div(style = "height: 0px; opacity: 0; background-color: lightblue",
           id = "two",
           "Next Div"),
  
  # create buttons to change the divs
  actionButton("btn-one", "Change Div", class = "btn"),
  actionButton("btn-two", "Change Div Back")
)

server <- function(input, output) {
}

shinyApp(ui, server)
```

Contents of `d3.js`
```{d3, include = F}
// define height of original div
var h = d3.select("#one").node().getBoundingClientRect().height;
var trans = 1000; // define transition time

// select first div element and make it disappear
d3.select('#btn-one').on("click", () => {
		d3.select('#one').transition().
		duration(trans).
		style("height", 0 + 'px')
		.style("opacity", 0);
		
// Select other div and replace 
		d3.select('#two').
		transition().
		duration(trans).
		style("height", h + 'px').
		style("opacity", 1);
});

// repeat same operation in reverse order
d3.select('#btn-two').on("click", () => {
		d3.select('#one').
		transition().
		duration(trans).
		style("height", h + 'px')
		.style("opacity", 1);
		
		d3.select('#two').
		transition().
		duration(trans).
		style("height", 0 + 'px').
		style("opacity", 0);
});
```

